Standard Candle Cosmology With Incomplete Spectroscopic Classification

### 2015-10-20 doc/article.tex
New version

\section{DES SN Survey Model}
\label{snIamodel:sec}
\subsection{Model}
This Section presents a candidate model
for supernova surveys that has the fidelity needed for the photometric
supernova analysis of the Dark Energy Survey and the Large Synoptic Survey Telescope.
%SNe~Ia are not perfect standard candles but rather are calibrated candles whose
%peak brightnesses are a
%function of their internal properties discerned from their colors and light-curve evolution.
%Measurement uncertainty, particularly from photometric flux
%calibration and galaxy photometric redshifts, can play an important role in the statistical
%reach of experiments. 
The notation for the likelihood presents the data and fundamental model parameters
\begin{equation}
p(n, {{T}}_S,{{z}}_S, \theta_{S}, \theta_G |  \Omega_M, w, r, \alpha_{Ia}, \beta_{Ia},\sigma_{Ia}, \alpha_{\mathit{non-Ia}},\beta_{\mathit{non-Ia}}, \sigma_{\mathit{non-Ia}},  \sigma_\beta, Z, g, g_N, f_g),
\label{likelihood2:eqn}
\end{equation}
whose relations are sketched in the probabilistic graphical model in Figure~\ref{pgm:fig}.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=6.5in]{/Users/akim/project/abc/results/hdpgm.pdf} 
   \caption{Probabilistic Graphical Model for the SN~Ia analysis.  
   The forward modeling
   flow goes from left to right, starting with the description of the universe, observatory,
   and data.    A transient has host galaxy $g_i$, which determines its redshift $z_i$
   and galaxy parameters $\theta_{gi}$.
   A model with parameters $\theta_\mu$ fix the distance modulus $\mu_i$.
   The transient type $T$ depends on the host-galaxy parameters  $\theta_{gi}$
   and rates $r$.   The transient
   parameters $\theta_T^X$, $\theta_{Ti}^X$ (and perhaps the host galaxy) determine the luminosity $L$.       The 
   incoming photon flux $n$, $n_g$  are then fixed
   with redshift and distance modulus.
   The instrumental transmission function $\phi$ is calibrated with data ${Z}$ andflux
   gives the expected
   counts $\overline{f}$, $\overline{f}_g$. 
   The transient has realized light curves ($n$) and passes selection criteria
   through $\tau_i$ to be included in the sample.  Some transients have spectral data
   ${X}_S$.  The properties of the galaxy identified as the host galaxy is $\theta_G$. 
   \label{pgm:fig}}
\end{figure}

The measurements are
\begin{itemize}
\item $n$: the measured counts for each photometric observation;
\item ${{T}}_S$, ${{z}}_S$, $\theta_S$: the spectroscopic type, redshift, and transient parameters available for a subset of transients;
\item $\theta_{G1}$, $\theta_{G2}$: the spectroscopic redshifts of the galaxies identified
as host and neighbor from imaging.
\end{itemize}
The model parameters are
\begin{itemize}
\item $\Omega_M$, $w$: the mass density and dark-energy equation of state parameters of a flat,
constant-$w$ dark-energy cosmology;
\item $r_1$, $r_2$:  the parameters
that describe the redshift-dependent relative fraction of SNe~Ia and non-Ia's;
\item $\alpha_{Ia}$, $\sigma_{Ia}$, $\alpha_{\mathit{non-Ia}}$, $\sigma_{\mathit{non-Ia}}$:
the luminosity (times $4\pi$) and intrinsic dispersion for SNe~Ia and non-Ia at peak brightness;
\item $\beta_{Ia}$, $\beta_{non-Ia}$: the parameters for the internal diversity within SNe~Ia and
non-Ia's respectively, e.g.\ SALT2's  $\alpha$ and $\beta$;
\item $\sigma_\beta$: parameters that describe the distribution of intrinsic supernova
parameters of SNe~Ia and non-Ia;
\item $Z$: the parameters that describe the transmission functions, i.e.\ photometric calibration;
\item $g$, $g_N$: the parameters that describe the candidate host and neighbor galaxies,
to be more expansive than the previous $z$ parameter;
\item $f_g$: the surface brightness of the background light at the location of the transient
needed for to extract the photometry of the point-source transient.
\end{itemize}
Not listed explicitly in Eqn.~\ref{likelihood:eqn} but shown in Figure~\ref{toypgm:fig} are latent model parameters for each transient:
\begin{itemize}
%\item $z$: the cosmological redshifts of the host and neighboring galaxies;
\item $T$: the class of the objects, SN~Ia or non-Ia;
\item $\theta_{Ia}$, $\theta_{non-Ia}$: the transient parameters that describe individual
transients, e.g.\ SALT2's $x_1$, $c$;
\item $L$: the luminosity of the transient.
\end{itemize}
There are  fixed parameters
\begin{itemize}
\item $\mu$: the distance modulus given the cosmological parameters and the redshift;
\item{$f$}: the flux of the transient given the luminosity, distance modulus, and
redshift;
\item{$\overline{n}_X$}: the expected counts of the transient and background light given
the flux and transmission.
\end{itemize}

The relationships between the parameters for the underlying population are now described.  What follows is an initial proposal
based on current supernova analyses, but many different models need to be
considered as the understanding of transients improve
and for probing systematics by comparing the
performance of competing models.

\begin{itemize}
\item Intrinsic transient parameters are dependent on their distributions and the host galaxy.
The SNe~Ia parameters are drawn as
\begin{equation}
\theta_{Ia} | \sigma_\beta, g \sim \mathcal{N}(0, \sigma_{\beta,Ia}),
\end{equation}
where $\theta_{Ia} =\{x_1, c\}$ are the SALT2 light-curve shape and color parameters.
A more complex model can be based on non-Normal distributions
and/or including SN properties directly tied to characteristics (e.g.\ mass, star formation
rate) of the host galaxy.

@@

%\begin{equation}
%\theta_{non-Ia} | \sigma_\beta, g  \sim \mathcal{N}(0, \sigma_{\beta, non-Ia}),
%\end{equation}
%where the parameter is the realization of an intrinsic magnitude dispersion.
%A more complex model can include multiple non-Ia populations each with its own
%intrinsic parameters.
\item Spectroscopically observed transient parameters depend on the intrinsic transient
parameters and type.  $\theta_{Si}$ may be ignored in a zero'th order analysis that
does not relate spectral features with SN~Ia absolute magnitudes.   If spectral
features are considered, the simple model
\begin{equation}
\theta_{Si} (\theta_{Ia}, \theta_{non-Ia}, T)=\theta_T
\end{equation}
assumes an unambiguous mapping between supernova parameters and spectral signal.
{\bf We need to decide what SN spectral parameters, if any, we will consider in our analysis.}
\item The type is drawn from the host-dependent relative rate between SNe~Ia and non-Ia's 
\begin{equation}
T | r, g \sim \rm{Bernoulli}(\pi(g,r)),
\end{equation}
where $\pi$ is a model for the relative rates of the different types as a
function of galaxy parameters and redshift.  {\bf A reasonable choice for $\pi$ needs to be
selected, any volunteers?}
\item Supernova type is assumed to be unambiguously encoded in the
spectrum:
\begin{equation}
T_S (T)  = T.
\end{equation}
\item The observed redshift is taken to be the host galaxy redshift
\begin{equation}
z_S (g_z) =g_z
\end{equation}
where $g_z$ is the element in $g$ that corresponds to redshift.
Added sophistication includes redshift
measurement uncertainty and random peculiar velocities. The study of coherent
flow requires added model complexity.
\item The underlying SED of transients are modeled.  For SNe~Ia the SALT2 model
is an option, although new models that separate dust and intrinsic color or
identify new SN~Ia parameters are expected in the near future.  Absolute magnitudes
that depend on host-galaxy properties need to be considered. 
For SNe~Ia $T=1$
\begin{equation}
L(t,\lambda) |  \alpha_{Ia}, \beta_{Ia}, \sigma_{Ia},  \theta_{Ia}, g, x_1, c \sim L_{SALT2}(t-t_0; \alpha_{Ia}, x_1, c) \gamma(g, \beta) \mathcal{N}(0,\sigma_{Ia}),
\end{equation}
where the first  term is the SALT2 light-curve and absolute magnitude
model with a realized error snake,
and $\gamma(g, \beta)$ a step function with free parameters in $\beta$ that
describe the step size.

{\bf The non-Ia model needs to be identified.}
\item The theoretical distance modulus
\begin{equation}
\mu(g_z; \Omega_M, w) = \mu_{wCDM}(g_z; \Omega_M, w)
\end{equation}
is the predicted distance modulus of a flat dark energy cosmology.
\item The flux incident on Earth is the redshifted and time-dilated SED scaled
by the distance modulus.
\begin{equation}
f(t,\lambda; L) = 10^{-\mu/2.5} L\left(t/(1+g_z),\lambda/(1+g_z)\right)
\end{equation}
\item The expected light curve in counts depends on the flux and throughput of the
optical chain.  A first model encapsulates flux calibration through a zeropoint
\begin{equation}
\overline{n}(f, Z) = 10^{Z/2.5} \int  f(\lambda) T_0(\lambda)d\lambda,
\end{equation}
where $T_0$ is a fiducial transmission function for the appropriate filter.  A more sophisticated
analysis can have a transmission function specific to each observation.
\item The measured surface brightness of the host galaxy depends on the underlying
surface brightness convolved with the point spread function and integrated
over the transmission
\begin{equation}
\overline{n}_g =    \int (f_g(\lambda, RA, Dec) \otimes PSF) T(\lambda;Z)d\lambda.
\end{equation}
\item The measurement $n$ is the superposition of transient and background
light $\overline{n} + \overline{n}_g$.
\end{itemize}

\subsection{Measurements}
\label{measurements:sec}
In the following discussion,
spectroscopic and galaxy measurements are assumed
to provide prefect determinations of transient type and parameters, and
galaxy redshifts and parameters. 
\begin{align}
p(T_S|T) &= \delta(T_S-T),\\
p(\theta_S|\theta_T) &= \delta(\theta_S-\theta_T),\\
p(\theta_G|g) &= \delta(\theta_G-g),\\
p(z_S|g_z) &= \delta(z_S-g_z)
\label{specz:eqn}
\end{align}
The working group is striving to get spectroscopic
redshifts of all potential hosts.  People who are interested in SN spectral features
or host-galaxy correlations (Chris, Rachel, Mat) can develop these models.

The measured counts depend on the expected counts and measurement uncertainty
$C_P$ as a Poisson (or considering the high sky background Normal) distribution
\begin{equation}
n \sim \mathcal{N} (\overline{n}+\overline{n}_g ,C_P).
\label{ADU:eqn}
\end{equation}
Practically, the photometric extraction is handled by independent software, which calculates
the pdf of $n_{\mathit{SN}}$, the realized contribution of supernova light,
replacing this and the previous two bullet points.  Dillon is providing this.
For the rest of the discussion, $n_{\mathit{SN}}$ is treated a measurement
directly connected to the expected mean signal of the supernova
$\overline{n}$
\begin{equation}
n_{SN} \sim \mathcal{N} (\overline{n} ,C_P).
\end{equation}

Prior on $Z$.
\subsection{Constructing the Likelihood}

The analysis sample satisfy certain photometric and no successful spectroscopy
({\it mis} Set)  and photometric
and successful spectroscopic classification ({\it obs} Set) selection criteria, which for convenience
probably should  be chosen to be stringent than the live-time decisions made during the survey.
Neither the photometric nor spectroscopic selection criteria have been defined.
Identifying good spectroscopic criteria needs to be done: the OzDES folks and Ryan are natural
candidates.

%The likelihood for the analysis sample is related to the underlying likelihood through
%\begin{multline}
%p(n, {{T}}_S,{{z}}_S, \theta_{S}, \theta_G |  \Omega_M, w, \theta_r, \alpha_{Ia}, \beta_{Ia},\sigma_{Ia}, \alpha_{\mathit{non-Ia}},\beta_{\mathit{non-Ia}}, \sigma_{\mathit{non-Ia}},  Z, g, g_N, f_g)\\
%= \frac{\epsilon(n, {{T}}_S,{{z}}_S, \theta_{S}, \theta_G) p_U(n, {{T}}_S,{{z}}_S, \theta_{S}, \theta_G | \ldots)}
%{\sum_{T_S} \int dn\, dz_S\, d\theta_S\, d\theta_G\, \epsilon(n, {{T}}_S,{{z}}_S, \theta_{S}, \theta_G) p_U(n, {{T}}_S,{{z}}_S, \theta_{S}, \theta_G | \ldots)},
%\end{multline} 
%where $p_U$ is the underlying likelihood  and $\epsilon$ is the efficiency for inclusion
%in the sample (more precisely subsample as spectroscopic
%selection should produce different efficiencies for the {\it obs} and {\it mis} sets.)

%The efficiencies for photometric and spectroscopic selection are taken to be independent.
%The efficiency  to pass photometric selection is denoted as $\epsilon_P$.

The following is a sketch of the likelihood calculation for an example 
photometric selection criterion, which should serve as a useful guide once the criteria are fixed.
The criterion is for $\ge N$ photometric points
above a signal-to-noise threshold (translating to a distinct limiting signal $n_{0i}$
for measurement $i$ of a given transient) given $M$ available measurements. 
The fail/pass status of the measurements are stored as 0/1 in an array $b$.
There are $\sum_{k=N}^M \binom{M}{k}$ 
distinct $b$'s that allow acceptance of the light curve.  In the notation $b_{ijk}$, $i$ indexes
the configuration, $j$ the transient, and $k$ the photometric point of transient $j$.

The transients are divided into
the following subsets:
{\it obs} are transients observed spectroscopically while active, from which $T_S=1$ are those typed
as SN~Ia and $T_S=0$ as non-Ia.  {\it mis} transients are not observed
spectroscopically when active.  The data from each subset have different forms for their likelihood.
The  {\it obs} sample constrains and specifies the parameter distributions of the 
parameters missing from the {\it mis} sample; both samples
are assumed to be drawn from the same underlying distribution.

\subsubsection{Likelihood for the {\it obs} Set}
The likelihood for the {\it obs} set is
\begin{multline}
p(n_{SN}, {{T}}_S,{{z}}_S, \theta_{S}, \theta_G |  \Omega_M, w, r, \alpha_{Ia}, \beta_{Ia},\sigma_{Ia}, \alpha_{\mathit{non-Ia}},\beta_{\mathit{non-Ia}}, \sigma_{\mathit{non-Ia}},  \sigma_\beta, Z, g, g_N, f_g)\\
= \sum_i \int_L dL\, p(n_{SN}, {{T}}_S,{{z}}_S, \theta_{S}, \theta_G , b_i|  L, \ldots)
 P(L|\ldots) \\
 \propto \sum_i  \epsilon(n_{SN}; b_{i}) \int_L dL\, p(n_{SN}, {{T}}_S,{{z}}_S, \theta_{S}, \theta_G|  L, \ldots)P(L|\ldots).
\end{multline} 
It is proportional to the truncated likelihood of the underlying sample, where the truncation
is achieved through $\epsilon(n_{SN}; b_{i})$, the efficiency that a light curve is 
selected in for a given configuration.  For the limiting flux criteria
\begin{equation}
\epsilon(n_{SN}; b_{i})= \prod_{jk}\mathcal{H}\left((-1)^{b_{ijk}+1}(n_{SN,jk}-n_{0,jk})\right),
\end{equation}
where $\mathcal{H}$ is the Heaviside step function; when $b_{ijk}=1$ the step in the efficiency
is up and for $b_{ijk}=0$ the step is down.

%The likelihood without the configuration condition for our measurements is
%\begin{multline}
%p(n_{SN}, {{T}}_S,{{z}}_S, \theta_{S}, \theta_G | L, \ldots) \\
%= \delta(z_S - g_z) \delta(\theta_S-\theta_T) \delta(\theta_G -g) \delta(T_S-T)  \mathcal{N}
%\left(n_{SN} -\overline{n}_j(L ,\ldots), C_P \right).
%\end{multline}. 
%

%
%The probability for failing (passing) the selection criterion for a single point is
%\begin{equation}
%P(b_{ij} | L, \ldots) = b_{ij} + (-1)^{b_{ij}}\int_{-\infty}^{ADU_0} dn_{SN,i}  \mathcal{N}\left(n_{SN} - \overline{n}_j(L ,\ldots), C_P \right),
%\end{equation}
%the (complementary) cumulative distribution function for a normal distribution
%with variable
%$n_0$, i.e.\ 2 times the error function.

For measurements described in \S\ref{measurements:sec}, the likelihood for the {\it obs} Set is
\begin{multline}
p(n_{SN}, {{T}}_S,{{z}}_S, \theta_{S}, \theta_G |  \Omega_M, w, r, \alpha_{Ia}, \beta_{Ia},\sigma_{Ia}, \alpha_{\mathit{non-Ia}},\beta_{\mathit{non-Ia}}, \sigma_{\mathit{non-Ia}},  \sigma_\beta, Z, g, g_N, f_g)\\
\propto \sum_i  \epsilon(n_{SN}; b_{i})  \int_L dL\, 
\mathcal{N}\left(n_{SN} - \overline{n}_j(L ,
g_z=z_S, \theta_T=\theta_S, g=\theta_G, T=T_S, \ldots), C_P \right)\\
P(L|g_z=z_S, \theta_T=\theta_S, g=\theta_G, T=T_S, \ldots),
\end{multline} 
where the latent parameters $T$, $\theta_T$ are not integrated over.
The normalization factor is parameter-dependent and is set by the condition
\begin{equation}
\sum_{T_S} \int dn_{SN}\,d{{z}}_S\,d\theta_{S}\,d\theta_G\,p(n_{SN}, {{T}}_S,{{z}}_S, \theta_{S}, \theta_G |  \ldots)=1.
\end{equation}

\subsubsection{Likelihood for the {\it mis} Set}
The likelihood for the {\it mis} set follow the same pattern as that of the
{\it obs} set except for not having to account for spectroscopic data.
\begin{multline}
p(n_{SN}, \theta_{G_1} |  \Omega_M, w, r, \alpha_{Ia}, \beta_{Ia},\sigma_{Ia}, \alpha_{\mathit{non-Ia}},\beta_{\mathit{non-Ia}}, \sigma_{\mathit{non-Ia}},  \sigma_\beta, Z, g, g_N, f_g) \\
 \propto \sum_i  \epsilon(n_{SN}; b_{i}) 
\int_L dL\, 
\sum_k 
\pi_k\mathcal{N}\left(n_{SN} - \overline{n}_j(L ,
g=\theta_{G_k}, \ldots), C_P \right)
P(L|g=\theta_{G_k}, \ldots), 
\end{multline} 
where $\pi_k$ is the probability that the $k$'th galaxy is the host, provided by Ravi.

>>>>

\section{DES SN Survey Model}
\label{snIamodel:sec}
\subsection{Model}
This Section presents a candidate model
for supernova surveys that has the fidelity needed for the photometric
supernova analysis of the Dark Energy Survey and the Large Synoptic Survey Telescope.
This model captures the current treatment in current analysis but expanded to include
photometry-only transients.  This may not be the ultimate model used, indeed many models
need to be considered for model testing.

%SNe~Ia are not perfect standard candles but rather are calibrated candles whose
%peak brightnesses are a
%function of their internal properties discerned from their colors and light-curve evolution.
%Measurement uncertainty, particularly from photometric flux
%calibration and galaxy photometric redshifts, can play an important role in the statistical
%reach of experiments. 
The notation for the likelihood presents the data and fundamental model parameters
\begin{multline}
\sum_T \int d\theta_{Ia}\, d\theta_{!Ia}\, dL\, p(n, {{T}}_S,{{z}}_S, \theta_{S}, G | T, \theta_{Ia},
\theta_{!Ia}, L, \mathcal{S}, \Omega_M, w, r, \Theta_{Ia}, \Theta_{\mathit{!Ia}}, Z, g) \\
p(L|T, \theta_{Ia},
\theta_{!Ia},  \mathcal{S}, \Omega_M, w, r, \Theta_{Ia},\Theta_{\mathit{!Ia}}, Z, g)
p(T, \theta_{Ia},
\theta_{!Ia}|  \mathcal{S}, \Omega_M, w, r, \Theta_{Ia},\Theta_{\mathit{!Ia}}, Z, g),
\label{likelihood2:eqn}
\end{multline}
whose relations are sketched in the probabilistic graphical model in Figure~\ref{pgm:fig}.
$\mathcal{S}$ indicates that this likelihood corresponds to the analysis sample. 
For shorthand we asign $A=\{n, {{T}}_S,{{z}}_S, \theta_{S}, G \}$, $B=\{T, \theta_{Ia},
\theta_{!Ia}, L\}$, and $C=\{\Omega_M, w, r, \Theta_{Ia}, \Theta_{\mathit{!Ia}}, Z, \}$
so that the likelihood is more compactly expressed as $\int dB\, p(AB| \mathcal{S}, C)p(B| \mathcal{S}, C)$.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=6.5in]{/Users/akim/project/abc/results/hdpgm.pdf} 
   \caption{Probabilistic Graphical Model for the SN~Ia analysis.  
   The forward modeling
   flow goes from left to right, starting with the description of the universe, observatory,
   and data.    A transient has host galaxy $g_i$, which determines its redshift $z_i$
   and galaxy parameters $\theta_{gi}$.
   A model with parameters $\theta_\mu$ fix the distance modulus $\mu_i$.
   The transient type $T$ depends on the host-galaxy parameters  $\theta_{gi}$
   and rates $r$.   The transient
   parameters $\theta_T^X$, $\theta_{Ti}^X$ (and perhaps the host galaxy) determine the luminosity $L$.       The 
   incoming photon flux $n$, $n_g$  are then fixed
   with redshift and distance modulus.
   The instrumental transmission function $\phi$ is calibrated with data ${Z}$ andflux
   gives the expected
   counts $\overline{f}$, $\overline{f}_g$. 
   The transient has realized light curves ($n$) and passes selection criteria
   through $\tau_i$ to be included in the sample.  Some transients have spectral data
   ${X}_S$.  The properties of the galaxy identified as the host galaxy is $G$. 
   \label{pgm:fig}}
\end{figure}

The measurements are
\begin{itemize}
\item $n$: the measured counts for each photometric observation for all transients in the sample;
\item ${{T}}_S$, ${{z}}_S$, $\theta_S$: the spectroscopic type, redshift, and transient parameters available for a subset of transients;
\item $G$: measured properties of the galaxy identified
as the host.
\end{itemize}
The top-level model parameters are
\begin{itemize}
\item $\Omega_M$, $w$: the mass density and dark-energy equation of state parameters of a flat,
constant-$w$ dark-energy cosmology;
\item $r$:  the parameters
that describe the redshift-dependent relative fraction of SNe~Ia and non-Ia's;
\item $\Theta_{Ia}$, $\Theta_{\mathit{!Ia}}$:  All the fundamental parameters used to describe
the SN~Ia and non-Ia populations respectively, such as the mean
absolute magnitude, intrinsic magnitude dispersion, the distribution of internal parameters
(i,e,\ the distributions of SALT2's $x_1$ and $c$), the relationship between internal
parameters and magnitude (i.e.\ SALT2's $\alpha$, $\beta$);
\item $Z$: the parameters that describe the transmission functions, i.e.\ photometric calibration;
\item $g$: the parameters that describe the candidate host and neighbor galaxies
including redshift $g_z$.
\end{itemize}
The second-level parameters are
\begin{itemize}
%\item $z$: the cosmological redshifts of the host and neighboring galaxies;
\item $T$: the class of the objects, SN~Ia or non-Ia;
\item $\theta_{Ia}$, $\theta_{!Ia}$: the transient parameters that describe individual
transients (i.e..\ SALT2's $x_1$, $c$);
\item $L(t,\lambda)$: the luminosity of the transient.
\end{itemize}
There are  fixed parameters
\begin{itemize}
\item $\mu$: the distance modulus given the cosmological parameters and the redshift;
\item{$f$}: the flux of the transient given the luminosity, distance modulus, and
redshift.
%\item{$\overline{n}$}: the expected counts of the transient given
%the flux and transmission.
\end{itemize}

The relationships between the parameters are now described:
\begin{itemize}
\item Intrinsic transient parameters are dependent on their distributions and the host galaxy.
The SNe~Ia parameters are drawn as
\begin{equation}
\theta_{Ia} | \Theta_{Ia}, g \sim \mathcal{N}(0, \Theta_{ia, \theta}),
\end{equation}
where $\theta_{Ia} =\{x_1, c\}$ are the SALT2 light-curve shape and color parameters
and $\Theta_{ia, \theta}$ describes the covariance of their distributions.
A more complex model can be based on non-Normal distributions
and/or including SN properties directly tied to characteristics (e.g.\ mass, star formation
rate) of the host galaxy.

@@

%\begin{equation}
%\theta_{non-Ia} | \sigma_\beta, g  \sim \mathcal{N}(0, \sigma_{\beta, non-Ia}),
%\end{equation}
%where the parameter is the realization of an intrinsic magnitude dispersion.
%A more complex model can include multiple non-Ia populations each with its own
%intrinsic parameters.
\item Spectroscopically observed transient parameters depend on the intrinsic transient
parameters and type.  $\theta_{Si}$ may be ignored in a zero'th order analysis that
does not relate spectral features with SN~Ia absolute magnitudes.   If spectral
features are considered, the simple model
\begin{equation}
\theta_{S} (\theta_{Ia}, \theta_{!Ia}, T)=\theta_{T}
\end{equation}
assumes an unambiguous mapping between supernova parameters and spectral signal.
{\bf We need to decide what SN spectral parameters, if any, we will consider in our analysis.}
\item The type is drawn from the host-dependent relative rate between SNe~Ia and non-Ia's 
\begin{equation}
T | r, g \sim \rm{Bernoulli}(\pi(r,g,T)),
\end{equation}
where $\pi$ is a model for the relative rate of SNe~Ia as a
function of galaxy parameters and redshift.  {\bf A reasonable choice for $\pi$ needs to be
selected.}
\item Supernova type is assumed to be unambiguously encoded in the
spectrum:
\begin{equation}
T_S (T)  = T.
\end{equation}
\item The observed redshift is taken to be the host galaxy redshift
\begin{equation}
z_S (g_z) =g_z
\end{equation}
where $g_z$ is the element in $g$ that corresponds to redshift.
Added sophistication includes redshift
measurement uncertainty and random peculiar velocities. The study of coherent
flow requires added model complexity.
\item The underlying SED of transients are modeled.  For SNe~Ia the SALT2 model
is an option, although new models that separate dust and intrinsic color or
identify new SN~Ia parameters are expected in the near future.  Absolute magnitudes
that depend on host-galaxy properties need to be considered. 
\begin{equation}
L(t,\lambda) |  \Theta_{Ia}, \theta_{Ia}, g \sim L_{SALT2}(t-t_0; \alpha, \beta, x_1, c) \gamma(g, \Theta_{Ia,g}) \mathcal{N}(0,\sigma_{Ia}),
\end{equation}
where the first  term is the SALT2 light-curve and absolute magnitude
model with a realized error snake,
and $\gamma(g, \Theta_{Ia})$ a step function to model the ``mass step''.  Recall that the  fundamental SALT2 parameters 
$\alpha$, $\beta$, and intrinsic dispersion $\sigma_{Ia}$, and the parameters that
relate SN magnitude and host,
are included in $\Theta_{Ia}$; the parameters in the SALT2 fit $t_0$, $x_1$, $c$
are included in $\theta_{Ia}$.

{\bf The non-Ia model needs to be identified.  Perhaps PSNID.}
\item The theoretical distance modulus
\begin{equation}
\mu(g_z; \Omega_M, w) = \mu_{wCDM}(g_z; \Omega_M, w)
\end{equation}
is the predicted distance modulus of a flat dark energy cosmology.
\item The flux incident on Earth is the redshifted and time-dilated SED scaled
by the distance modulus.
\begin{equation}
f(t,\lambda; L) = 10^{-\mu/2.5} L\left(t/(1+g_z),\lambda/(1+g_z)\right)
\end{equation}
\item The expected light curve in counts depends on the flux and throughput of the
optical chain.  A first model encapsulates flux calibration through a zeropoint
\begin{equation}
n(f, Z) = 10^{Z/2.5} \int  f(\lambda) T_0(\lambda)d\lambda,
\end{equation}
where $T_0$ is a fiducial transmission function for the appropriate filter.  A more sophisticated
analysis can have a transmission function specific to each observation.
\item The host-galaxy information is directly encoded in the measurements
\begin{equation}
G(g)=g.
\end{equation}
\end{itemize}

\subsection{Measurements}
\label{measurements:sec}
Spectroscopic measurements are assumed
to provide prefect determinations of transient type and parameters, and
galaxy redshifts and parameters. 
\begin{align}
p(T_S|T) &= \delta(T_S-T),\\
p(\theta_S|\theta_T) &= \delta(\theta_S-\theta_T),\\
p(z_S|g_z) &= \delta(z_S-g_z).
\label{specz:eqn}
\end{align}
People who are interested in SN spectral features
or host-galaxy correlations (Chris, Rachel, Mat) can develop these models.

For the photometric sample, there is uncertainty in the identity of the host galaxy.
From a set of potential host galaxies $\hat{G}$ each with probability $\pi$ of being the correct host,
\begin{equation}
G \sim \prod_i \delta(G-\hat{G}_i) \pi_i.
\end{equation}
Ravi can provide for each transient a list of candidate hosts and their probabilities.
The working group is striving to get spectroscopic
redshifts of all potential hosts, which is included in $G$.
For the spectroscopic sample the host is unambigous
\begin{equation}
G = g.
\end{equation}

Practically the photometric extraction is handled by independent software, which calculates
the $n$ with uncertainties $C_P$
by modeling and separating out polluting background light from the transient
signal
\begin{equation}
n \sim \mathcal{N}\left(10^{Z/2.5} \int  f(\lambda;A, B) T_0(\lambda)d\lambda,C_P\right).
\end{equation}
Dillon is providing this.
Note that the model has explicit calibration parameters that are to be fit.  An alternative
approach is to include calibration uncertainties in $C_P$.

\subsection{Sample Selection}
The transients that enter the analysis satisfy some selection criteria.  Trivially, they
have to be discovered in the first place and often more stringent photometric signal-to-noise
requirements are applied.  Additionally, the probability for a transient to have successful
spectroscopy may depend on parameters.  These criteria have not yet been determined/selected
(Argonne for photometry, OzDES/Ryan for spectroscopy are candidates to do this);
presented here is a sketch of the likelihood calculation for a reasonable
photometric selection criterion.
For each transient $\nu$, the criterion is to have $\ge N$ photometric points
above a signal-to-noise threshold there  given $M_\nu$ available measurements:
the signal-to-noise threshold corresponding to a flux threshold
 $n_{0,\nu i}$.

Call the likelihood of the underlying distribution without sample selection $P(A|C)$.
Let the array $b_{\nu}$ be an array that contains $0$ or $1$ based on
whether the corresponding photometric point is below or passes threshold. 
Let the larger array $b$ be a combination of the $b_\nu$ from all transients.
There are $2^{\sum_\nu M_\nu}$ possible permutations of $b$, and the underlying
likelihood can be broken up into the sum of all possible combinations
\begin{align}
p(A|C) &= \sum_k p(A, b_k| C)\\
& = \sum_k \epsilon(n, b_k)p(A| C),
\end{align}
where
\begin{equation}
\epsilon(n; b_{k})= \prod_{\nu i}\mathcal{H}\left((-1)^{b_{k\nu i}+1}(n_{k\nu i}-n_{0,\nu i})\right),
\end{equation}
where $\mathcal{H}$ is the Heaviside step function; when $b_{k\nu i}=1/0$ the step in the efficiency
is up/down.  The indices correspond to configuration $k$,
transient $\nu$, and photometric point $i$.

For the analysis sample, are a smaller number of combinations of $b$ that satisfy
the selection criteria;
instead of $2^{M_\nu}$ there are $\sum_{k=N}^{M_{\nu}} \binom{M_{\nu}}{k}$ 
distinct $b_{\nu}$'s that allow acceptance of transient $\nu$.  The selection efficiency
for the complement set of possible $b$'s is zero.
The sample likelihood is then as
\begin{equation}
p(A|\mathcal{S}, C) = \frac{\sum_k \epsilon(n, b_k)p(A| C)}{\int dA\, \sum_k \epsilon(n, b_k)p(A| C)},
\end{equation}
where the sum is
over only the configurations that meet the selection criteria.

The measurement uncertainties in the light curves $n$ are independent of the other
measurements, and the measurement uncertainties are drawn from a Normal distribution:
the likelihood is assembled from truncated Gaussians with the integrals over $n$ in the numerator
proportional to error functions.

\subsection{Likelihoods}
The transients are divided into
the following subsets:
{\it obs} are transients observed spectroscopically while active, from which $T_S=1$ are those typed
as SN~Ia and $T_S=0$ as non-Ia.  {\it mis} transients are not observed
spectroscopically when active.  The data from each subset have different forms for their likelihood.
The  {\it obs} sample constrains and specifies the parameter distributions of the 
parameters missing from the {\it mis} sample.

\subsubsection{Likelihood for the {\it obs} Set}
The likelihood for the {\it obs} set is
\begin{multline}
p(A|\mathcal{S},C) \propto
\sum_k \epsilon(n, b_k) 
\sum_T \int d\theta_{Ia}\, d\theta_{!Ia}\, dL\, p(n, {{T}}_S,{{z}}_S, \theta_{S}, G | T,  L, \Omega_M, w, r, \Theta_{Ia}, \Theta_{\mathit{!Ia}}, Z, g) \\
p(L| T, \theta_{Ia},
\theta_{!Ia},  \Omega_M, w, r, \Theta_{Ia},\Theta_{\mathit{!Ia}}, Z, g)
p(T, \theta_{Ia}
\theta_{!Ia}|  \Omega_M, w, r, \Theta_{Ia},\Theta_{\mathit{!Ia}}, Z, g).
\end{multline}

For our model and measurements
\begin{multline}
p(A|\mathcal{S},C) \propto 
\sum_k \epsilon(n, b_k) \int dL\,
p(n |  T= T_S, \theta_{T}=\theta_S, g=G, L, C)
p(L| T, \theta_{T_S}, r, \Theta_{T_S}, g)\\
\pi(r, z_S, T_S)p( \theta_{T_S}|  \Theta_{T_S},g=G) .
\end{multline}

\subsubsection{Likelihood for the {\it mis} Set}
The likelihood for the {\it mis} set follows the same pattern as that of the
{\it obs} set except for not having to account for spectroscopic data
\begin{multline}
p(n, \theta_{G} |  \mathcal{S}, C) 
 \propto \sum_k \epsilon(n; b_{k}) 
\sum_T \int d\theta_{Ia}\, d\theta_{!Ia}\, dL\, 
\sum_l
\pi_l
p(n |  g=\hat{G}_l, B, C)
p(L| T, \theta_{Ia}, \theta_{!Ia}, r, \Theta_{Ia}, \Theta_{!Ia}, g)\\
\pi(r, \hat{G}_l, T)p( \theta_{Ia},
\theta_{!Ia}|  \Theta_{Ia},\Theta_{\mathit{!Ia}}, g=\hat{G}_l).
\end{multline}

%\begin{multline}
%p(n_{SN}, G |  \Omega_M, w, \theta_r, \alpha_{Ia}, \beta_{Ia},\sigma_{Ia}, \alpha_{\mathit{non-Ia}},\beta_{\mathit{non-Ia}}, \sigma_{\mathit{non-Ia}},  Z, g, g_N, f_g)\\
%= \sum_i \int_L dL\, p(n_{SN}, G | b_i, L, \ldots)
%P(b_{i} |L, \ldots) P(L|\ldots) ,
%\end{multline} 
%\begin{equation}
%p(n_{SN}, G | b_i, L, \ldots)
%  =  \frac{ \prod_j
% \epsilon(n_{SN,j}; b_{ij})
%  p(n_{SN}, G | L, \ldots)}
%  {\prod_j P(b_{ij}{ |L,\dots)}}.
%\label{misconf:eqn}
%\end{equation}
